### 2022-11-14 submission/manuscript.Rmd
Initial commit

---
bibliography: references.bib
output:
  pdf_document:
    keep_tex: true
csl: asm.csl
geometry: margin=1.0in
header-includes:
 - \usepackage{booktabs}
 - \usepackage{longtable}
 - \usepackage{array}
 - \usepackage{multirow}
 - \usepackage{wrapfig}
 - \usepackage{float}
 - \usepackage{colortbl}
 - \usepackage{pdflscape}
 - \usepackage{tabu}
 - \usepackage{threeparttable}
 - \usepackage{threeparttablex}
 - \usepackage[normalem]{ulem}
 - \usepackage{makecell}
 - \usepackage{setspace}
 - \doublespacing
 - \usepackage[left]{lineno}
 - \usepackage{helvet} % Helvetica font
 - \renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
 - \usepackage[T1]{fontenc}
---

```{r knitr_settings, tidy=TRUE, eval=TRUE, echo=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("message" = FALSE)
opts_chunk$set("cache" = FALSE)

inline_hook <- function(x){
  print(x)
  if(is.list(x)){
    x <- unlist(x)
  }
  if(is.numeric(x)){
    if(abs(x - round(x)) < .Machine$double.eps^0.5){
      paste(format(x,big.mark=',', digits=0, scientific=FALSE))
    } else {
      paste(format(x,big.mark=',', digits=2, nsmall=2, scientific=FALSE))
    }
  } else {
      paste(x)
  }
}
knitr::knit_hooks$set(inline=inline_hook)
```

\vspace*{10mm}

# The *Riffomonas* YouTube Channel: An Educational Resource to Foster Reproducible Research Practices

\vspace{15mm}

Running title: *Riffomonas* YouTube Channel

\vspace{15mm}

Patrick D. Schloss${^1}$${^\dagger}$

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed: \href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI 48109

\vspace{35mm}

### Educational resource

\newpage
\linenumbers

## Abstract
Limit the abstract to 50 words or fewer

\newpage

Limit the paper to 500 words or fewer

[@Schloss2018a]
[@Schloss2018b]

Code Club

https://www.youtube.com/c/RiffomonasProject

Need for all scientists to strengthen their data science skills
* Focus on reproducible research methods
* command line, R, Rmarkdown, version control, data visualization, project organization

Previously...
* Developed Riffomonas reproducible research tutorial series
* Developed other online tutorials teaching scientists to use in R with microbiome (minimalR) and more general (generalR) data

Current state of the art...
* Workshop-based tutorials - intensive learning opportunities (minimalR/generalR and Carpentries)
* Books - use toy datasets, highlight individual commands

The *Riffomonas* YouTube Channel...
* Repository for Reproducible Research Tutorial series
* Code Club series

Code Club
* Born out of pandemic and need for community and desire to fill the hole in offerings for repeated practice, applying concepts in different contexts, and intgration of concepts in project-based approach
* So far...
  - Live coded the process of developing, writing, and publishing a paper
  - Use "real" and not toy datasets - microbiome, weather, commodity prices, COVID-19 vaccine attitudes - to develop concepts
* Posted 1-3 videos a week
* Recently started a free weekly email newsletter with practice problems that parallel content in the videos

Availability of code through Jekyll powered blog and repositories within a GitHub-based project

Pedagogy...
* Offering encouragement
* Normalize failure
* Repeated practice opportunities
* Ability to apply concepts in different contexts
* Integration of concepts in a project-based approach

## Acknowledgements

\newpage

## References

<div id="refs"></div>

\newpage

**Figure 1. Lorem ipsum dolor sit amet.** Consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.