### 2022-11-14 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\vspace*{10mm}

\hypertarget{the-riffomonas-youtube-channel-an-educational-resource-to-foster-reproducible-research-practices}{%
\section{\texorpdfstring{The \emph{Riffomonas} YouTube Channel: An
Educational Resource to Foster Reproducible Research
Practices}{The Riffomonas YouTube Channel: An Educational Resource to Foster Reproducible Research Practices}}\label{the-riffomonas-youtube-channel-an-educational-resource-to-foster-reproducible-research-practices}}

\vspace{15mm}

Running title: \emph{Riffomonas} YouTube Channel

\vspace{15mm}

Patrick D. Schloss\({^1}\)\({^\dagger}\)

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

\(1\) Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI 48109

\vspace{35mm}

\hypertarget{educational-resource}{%
\subsubsection{Educational resource}\label{educational-resource}}

\newpage
\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Limit the abstract to 50 words or fewer

\newpage

Limit the paper to 500 words or fewer

(1) (2)

Need for all scientists to strengthen their data science skills * Focus
on reproducible research methods * command line, R, Rmarkdown, version
control, data visualization, project organization

Previously\ldots{} * Developed Riffomonas reproducible research tutorial
series * Developed other online tutorials teaching scientists to use in
R with microbiome (minimalR) and more general (generalR) data

Current state of the art\ldots{} * Workshop-based tutorials - intensive
learning opportunities (minimalR/generalR and Carpentries) * Books - use
toy datasets, highlight individual commands

The \emph{Riffomonas} YouTube Channel\ldots{} * Repository for
Reproducible Research Tutorial series * Code Club series

Code Club * Born out of pandemic and need for community and desire to
fill the hole in offerings for repeated practice, applying concepts in
different contexts, and intgration of concepts in project-based approach
* So far\ldots{} - Live coded the process of developing, writing, and
publishing a paper - Use ``real'' and not toy datasets - microbiome,
weather, commodity prices, COVID-19 vaccine attitudes - to develop
concepts * Posted 1-3 videos a week * Recently started a free weekly
email newsletter with practice problems that parallel content in the
videos

Availability of code through Jekyll powered blog and repositories within
a GitHub-based project

Pedagogy\ldots{} * Offering encouragement * Normalize failure * Repeated
practice opportunities * Ability to apply concepts in different contexts
* Integration of concepts in a project-based approach

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{1}
\leavevmode\vadjust pre{\hypertarget{ref-Schloss2018a}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Schloss PD}. 2018. The {R}iffomonas reproducible
research tutorial series. Journal of Open Source Education
\textbf{1}:13.
doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}.}

\leavevmode\vadjust pre{\hypertarget{ref-Schloss2018b}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{\textbf{Schloss PD}. 2018. Identifying and overcoming
threats to reproducibility, replicability, robustness, and
generalizability in microbiome research. {mBio} \textbf{9}.
doi:\href{https://doi.org/10.1128/mbio.00525-18}{10.1128/mbio.00525-18}.}

\end{CSLReferences}

\newpage

\textbf{Figure 1. Lorem ipsum dolor sit amet.} Consectetur adipisicing
elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi
ut aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa
qui officia deserunt mollit anim id est laborum.

\end{document}